---
title: "dataanalysis_exercise"
editor: visual
---

### Description of data

The data I chose for this exercise is data from the CDC that contains the funds the CDC received from the federal government in a form of five differnt congressional acts: the Coronavirus Preparedness and Response Supplemental Appropriations Act, 2020; Coronavirus Aid, Relief, and Economic Security Act; Paycheck Protection Program and Health Care Enhancement Act; Coronavirus Response and Relief Supplemental Appropriations Act, 2021; and American Rescue Plan Act of 2021.

###Loadpackages 
```{r}
library(readr) #for loading csv files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(stats) #used for cleaning 
library(here) #to set paths
```

###Loaddata
#note the use of the here() package and not absolute paths
#notice here::here is to indicate package::function()
```{r}
data_location <- here::here("dataanalysis-exercise", "data","raw data",
                                          "CDC_COVID_State_Funding.csv")
rawdata <- readr::read_csv(data_location)
```

###explore the data 
```{r}
dplyr::glimpse(rawdata)
summary(rawdata)
head(rawdata)
skimr::skim(rawdata)
```

###cleandata1 
#the only variable that appears to have a lot of missing values is Sub-jurisdiction.
#I am going to extract the observations that have missing sub-jurisdictions. 
```{r}
df1 <- stats::na.omit(rawdata)
```

###cleandata2
#there are some funding amounts that are $0.
#I am only interested in funding that is over $0.
#to clean this up I will remove all the rows containing $0.
```{r}
df2 <- df1 %>% dplyr::filter(Amount != 0)
```

###cleandata3
#I am only interested in the supplemental act name.
#the following code will remove the topical area column and the award name column.
```{r}
df3 <- dplyr::select(df2,-c("Topical Area", "Award Name"))
```

###Variables of interest 
#Interest which Jurisdition has the most funding (amount).
#Also curious which supplemental act has the most funding.
