---
title: "exploration"
output:
  html_document:
    toc: FALSE
---

### Library for exploration
```{r}
library(tidyverse)
library(ggplot2)
library(Hmisc) #needed for the describe() function
library(dplyr)
```

### Getting a visual of the distribution of body temp
```{r}
Bodytemp_plot = ggplot(rawdata3,aes(x=BodyTemp)) +
  geom_density(color="darkblue", fill="lightblue")
Bodytemp_plot
bodytemp_summary
#there is a skew to the right with a mean at 98.94 
```

### Summary statistics for binary (yes or no) predictor variables
-   Nausea, Swollen lymph nodes and Insomia

```{r}
#I chose to use the describe function because it provides count (n), and the frequency and proportion of response
nausea_summary = describe(rawdata3$Nausea)
lymph_summary = describe(rawdata3$SwollenLymphNodes)
insomnia_summary = describe(rawdata3$Insomnia)
```

# Checking out the summaries of the binary predictor variables 
```{r}
nausea_summary
lymph_summary
insomnia_summary
```

All three binary variables appear to be pretty split, with no overwhelming majority. It is important to note that the majority of responses to the nausea question were no (65.1%). Also, the majority of the responses to the swollen lymph nodes question was no (57.3%). It is interesting to see that the majority of response to the insomnia question was yes (56.8%).

