---
title: "Model Evaluation"
editor: visual
output:
  html_document:
    toc: FALSE
---

### Library

```{r}
library(tidymodels)
library(here)
library(rsample)
library(parsnip) #building a model specification
library(yardstick)
library(dplyr)
```

### Loading the data

```{r}
location = here("Fluanalysis_exercise", "data", "cleandata", "cleandata.RDS")
clean_data = readRDS(location)
```

## Data Splitting

### We can use the rsample package to create an object that contains the information on how to split the data

```{r}
# Fix the random numbers by setting the seed 
# This enables the analysis to be reproducible when random numbers are used 
set.seed(222)

# Put 3/4 of the data into the training set 
data_split = initial_split(clean_data, prop = 3/4)
```

### Create data frames for the two sets

```{r}
train_data = training(data_split)
test_data  = testing(data_split)
```

## Creating a simple recipe that fits a logistic model to our categorical outcome of interest(Nausea), using all predictors.

### Initiating a new recipe

```{r}
flu_rec <- 
  recipe(Nausea ~ ., data = train_data) # . indicates all predictors are present

```

### Set a model

```{r}
lr_mod <- 
  logistic_reg() %>% 
  set_engine("glm")
```

### Using the workflow() function to create a simple workflow that fits a logistic model to all predictors using the glm engine

```{r}
flu_wflow <- 
  workflow() %>% 
  add_model(lr_mod) %>% 
  add_recipe(flu_rec)
flu_wflow
```

### A single function that can be used to prepare the recipe and train the model from the resulting predictors

```{r}
# fitting the model to the training data
flu_fit1 <- 
  flu_wflow %>% 
  fit(data = train_data)

# fitting the model to the test data 
flu_fit2 <- 
  flu_wflow %>% 
  fit(data = test_data)

```

### Extracting the model or recipe objects from the workflow in order to view model coefficients

```{r}
# Extracting coefficients for model fit to the training data
flu_fit1 %>% 
  extract_fit_parsnip() %>% 
  tidy() #tidy() function used to get a tidy tibble of the model coefficients 

#Extracting the coefficients for the model fit to the test data 
flu_fit2 %>% 
  extract_fit_parsnip() %>% 
  tidy() 
```

### Using augment() to predict the and save the predicted data with the dataset

```{r}
flu_aug1 = augment(flu_fit1, train_data)

flu_aug2 = augment(flu_fit2, test_data)

flu_aug1
flu_aug2
```

### Creating an ROC curve and producing estimates under the curve for the training data for predicted class probability of No

```{r}
flu_aug1 %>% 
  roc_curve(truth = Nausea, .pred_No)  %>% 
  autoplot

flu_aug1 %>% 
  roc_auc(truth = Nausea, .pred_No)
```

### Creating an ROC curve and producing estimates under the curve for the test data for predicted class probability of No

```{r}
flu_aug2 %>% 
  roc_curve(truth = Nausea, .pred_No)  %>% 
  autoplot

flu_aug2 %>% 
  roc_auc(truth = Nausea, .pred_No)
```

# Alternate Model

### Setting up the recipe for a model between main predictor and categorical outcome

```{r}
flu2_rec <- 
  recipe(Nausea ~ RunnyNose, data = train_data) # . indicates all predictors are present
```

### Set a model

```{r}
lr_mod2 <- 
  logistic_reg() %>% 
  set_engine("glm")
```

### Using the workflow() function to create a simple workflow that fits a logistic model to all predictors using the glm engine

```{r}
flu_wflow2 <- 
  workflow() %>% 
  add_model(lr_mod) %>% 
  add_recipe(flu_rec)
flu_wflow2
```

### A single function that can be used to prepare the recipe and train the model from the resulting predictors

```{r}
# fitting the model to the training data
flu_fit3 <- 
  flu_wflow %>% 
  fit(data = train_data)

# fitting the model to the test data 
flu_fit4 <- 
  flu_wflow %>% 
  fit(data = test_data)

```

### Extracting the model or recipe objects from the workflow in order to view model coefficients

```{r}
# Extracting coefficients for model fit to the training data
flu_fit3 %>% 
  extract_fit_parsnip() %>% 
  tidy() #tidy() function used to get a tidy tibble of the model coefficients 

#Extracting the coefficients for the model fit to the test data 
flu_fit4 %>% 
  extract_fit_parsnip() %>% 
  tidy() 
```

### Using augment() to predict the and save the predicted data with the dataset

```{r}
flu_aug3 = augment(flu_fit3, train_data)

flu_aug4 = augment(flu_fit4, test_data)

flu_aug3
flu_aug4
```

### Creating an ROC curve and producing estimates under the curve for the training data

```{r}
flu_aug3 %>% 
  roc_curve(truth = Nausea, .pred_No)  %>% 
  autoplot

flu_aug3 %>% 
  roc_auc(truth = Nausea, .pred_No)
```

### Creating an ROC curve and producing estimates under the curve for the test data

```{r}
flu_aug4 %>% 
  roc_curve(truth = Nausea, .pred_No)  %>% 
  autoplot

flu_aug4 %>% 
  roc_auc(truth = Nausea, .pred_No)
```



## Deanna Lanier Additions ##

*Instructions*
Add code that repeats the above steps, but now fits linear models to the continuous outcome. One model with all predictors, one model with just the main predictor. For that, you need to change your metric. RMSE is a good one to choose. You should find that a lot of the code your classmate wrote can copied/pasted/re-used with only minimal modifications. Make sure to add comments/documentation.

## add packages (only ones but listed above)
```{r}
## add library to complete RSME
library(Metrics)
```

### Loading the cleaned data

```{r}
location = here("Fluanalysis_exercise", "data", "cleandata", "cleandata.RDS")
data = readRDS(location)
```

### Generating training and test data sets

```{r}
# set seed. This helps with reproducibility 
set.seed(5678)
data_split=initial_split(data,prop = 3/4)

#create training and test
data_train=training(data_split)
data_test=testing(data_split)
```

### Generating a workflow with the continuous outcome with all data

```{r}
#recipe
recipe_1 = recipe(BodyTemp ~ ., data = data)

# initialize the model 
## I am using a linear model because of the continuous outcome
model_1 =  linear_reg() %>%
             set_engine("lm")

# create workflow
workflow_1 = 
             workflow() %>%
             add_model(model_1) %>%
             add_recipe(recipe_1)
workflow_1
```
### train the model with the training set
```{r}
model_train_1 = workflow_1 %>%
                fit(data = data_train)

model_train_1 %>%
  extract_fit_parsnip()%>%
  tidy()
```

## Predict 
```{r}
predict(model_train_1, data_test)
```

### Augment and RMSE
```{r}
#augment
model1_aug = augment(model_train_1, data_test)

model1_aug %>%
  select(BodyTemp, .pred)

#rmse
Metrics::rmse(actual = model1_aug$BodyTemp, predicted = model1_aug$.pred)
```

### Do the same with primary predictor 

```{r}
#recipe 
recipe_2 = recipe(BodyTemp ~ RunnyNose, data = data)

#workflow
workflow_2 =
             workflow() %>%
             add_model(model_1) %>%
             add_recipe(recipe_2)

#model fit train
model_2 = workflow_2 %>%
                fit(data = data_train)
```

```{r}
## output
model_2 %>%
  extract_fit_parsnip() %>%
  tidy()
```
### Predict
```{r}
predict(model_2, data_test)
```
```{r}
# augment
model_2_aug <- augment(model_2, data_test)

model_2_aug %>%
  select(BodyTemp, .pred)

#RSME
Metrics::rmse(actual = model_2_aug$BodyTemp, predicted = model_2_aug$.pred)
```
