---
title: "R Data Analysis Exercise"
output: 
  html_document:
      toc: FALSE
---

### Description of data

The data I chose for this exercise is data from the CDC that contains the funds the CDC received from the federal government in a form of five different congressional acts: the Corona virus Preparedness and Response Supplemental Appropriations Act, 2020; Corona virus Aid, Relief, and Economic Security Act; Paycheck Protection Program and Health Care Enhancement Act; Corona virus Response and Relief Supplemental Appropriations Act, 2021; and American Rescue Plan Act of 2021.

Load packages

```{r}
library(readr) #for loading csv files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(stats) #used for cleaning 
library(here) #to set paths
```

Load data

-   Note the use of the here() package and not absolute paths

-   Notice here::here is to indicate package::function()

```{r}
data_location <- here::here("dataanalysis-exercise", "data","raw data",
                                          "CDC_COVID_State_Funding.csv")
rawdata <- readr::read_csv(data_location)
```

Explore the Data

```{r}
dplyr::glimpse(rawdata)
summary(rawdata)
head(rawdata)
skimr::skim(rawdata)
```

Clean Data

-    the only variable that appears to have a lot of missing values is Sub-jurisdiction.

-   I am going to extract the observations that have missing sub-jurisdictions.

```{r}
df1 <- stats::na.omit(rawdata)
```

Clean Data Further

-    There are some funding amounts that are \$0.

-   I am only interested in funding that is over \$0.

-   To clean this up I will remove all the rows containing \$0.

```{r}
df2 <- df1 %>% dplyr::filter(Amount != 0)
```

Clean Data3 Even Further

-   I am only interested in the supplemental act name. The following code will remove the topical area column and the award name column.

```{r}
df3 <- dplyr::select(df2,-c("Topical Area", "Award Name"))
```

Variables of interest

-   Interested in which Jurisdiction has the most funding (amount).

-   Also curious which supplemental act has the most funding.

-----------------------------------------------Aidan Troha's Edit-------------------------------------------- Additional Packages

```{r}
library(ggplot2)
```

Selecting Data

-   First, we summarize the data based on location (Jurisdiction).

-   Also, we wish to convert the "Amount" column to one that can be easily readable by viewers.

-   We can achieve this by using the group-by(), summarise(), and mutate() functions.

```{r}
test1 <- df3 %>%
      group_by(Jurisdiction) %>%
      summarise(Funding = sum(Amount)) %>%
      mutate(`Funding (in Millions)` = Funding/1000000)
```

Plotting

-   We want to show the relationships between the state and the amount of funding received. Since there is such a disparity between the states with the most funding and the states with the least funding, I will only show the states with the highest level of funding.

```{r}
ggplot(mapping=aes(x=Jurisdiction,y=`Funding (in Millions)`)) +
       geom_col(data=slice_max(test1,order_by=`Funding (in Millions)`,n=5),
                aes(fill=c("red","yellow","green","blue","purple"))) +
       ylim(0,max(test1$`Funding (in Millions)`)) +
       theme(legend.position="none") +
       labs(title="States with the Highest Level of Funding")
```

Selecting data

-   As before, we summarise the data, but this time, we group by "Supplemental Act Name".

```{r}
test2 <- df3 %>%
  group_by(`Supplemental Act Name`) %>%
  summarise(Funding = sum(Amount)) %>%
  mutate(`Funding (in Millions)` = Funding/1000000)
```

Plotting2

-   We want to show the relationships between the cumulative amount of funding donated by Supplemental Act.

```{r}
ggplot(data=test2,mapping=aes(x=`Supplemental Act Name`,
                              y=`Funding (in Millions)`)) +
  geom_col(aes(fill=c("red","orange","yellow","green","blue","purple","black"))) +
  theme(legend.position="none") +
  scale_x_discrete(labels=scales::label_wrap(12)) +
  labs(title="Total Funding Awarded by Supplemental Act")
```
