---
title: "Tidy Tuesday Exercise"
output: 
  html_document:
    toc: FALSE
---

# Loading Packages

```{r}
library(tidytuesdayR)
library(tidyverse)
library(here)
library(Hmisc)
library(ggplot2)
library(plotly)
```

# Get the Data

```{r}
tuesdata <- tidytuesdayR::tt_load('2023-02-14')
tuesdata <- tidytuesdayR::tt_load(2023, week = 7)

age_gaps <- tuesdata$age_gaps
```

# Changing variable names for the older and younger character variables just so it's easy to read

```{r}
#changing actor gender variable
age_gaps <- age_gaps %>% 
  rename(older_actor_gender = character_1_gender)
age_gaps <- age_gaps %>% 
  rename(younger_actor_gender = character_2_gender)

#changing actor name variable name 
age_gaps <- age_gaps %>% 
  rename(older_actor_name = actor_1_name)
age_gaps <- age_gaps %>% 
  rename(younger_actor_name = actor_2_name)

#changing birthday actor name 
age_gaps <- age_gaps %>% 
  rename(older_actor_birthdate = actor_1_birthdate)
age_gaps <- age_gaps %>% 
  rename(younger_actor_birthdate = actor_2_birthdate)

#changing actor age variable
age_gaps <- age_gaps %>% 
  rename(older_actor_age = actor_1_age)
age_gaps <- age_gaps %>% 
  rename(younger_actor_age = actor_2_age)
```

# Using the describe function from the Hmisc package to see how many of the observations in the character_1\_gender variable are man and woman. Then producing a tibble with older actor variable

```{r}
#using the describe function from the Hmisc package to see how many of the observations in the character_1_gender variable are man and woman 
describe(age_gaps$older_actor_gender)

#producing tibble with older actor variable
freq_older_char <- age_gaps %>%
  group_by(older_actor_gender) %>%
  summarise(counts = n())
freq_older_char

```

# Producing plot for man vs woman that play the older character

```{r}
older_plot = ggplot(freq_older_char, aes(x = older_actor_gender, y = counts)) +
  geom_bar(fill = "#0073C2FF", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) + ggtitle("Count of older men characters vs older women characters")
older_plot
```

# Producing a tibble with frequencies of younger actor variable

```{r}
describe(age_gaps$younger_actor_gender)

freq_younger_char <- age_gaps %>%
  group_by(younger_actor_gender) %>%
  summarise(counts = n())
freq_younger_char
```

# Producing plot for man vs woman that play the younger character

```{r}
younger_plot = ggplot(freq_younger_char, aes(x = younger_actor_gender, y = counts)) +
  geom_bar(fill = "red", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) + ggtitle("Count of younger men characters vs younger women characters")
younger_plot
```

# Summary of findings so far

-   In the movies included in the data it is a more common for a male actor to play the older character when compared to women playing the older character

-   With that in mind, I am curious if there is any relationship between movies with an older man character and the release date of the movie

# Creating a data set with just men who are the older character

```{r}
year_vs_oldergender <- age_gaps %>% group_by(release_year, older_actor_gender)
year_vs_oldergender <-year_vs_oldergender %>% summarise(n = n())

#filtering just the men character from the women older characters 
year_vs_oldermen = filter(year_vs_oldergender, older_actor_gender == "man")

#removing the gender column because they are all male 
year_vs_oldermen <- year_vs_oldermen[,-2]
```

# Using ggplot2 to create a bar plot then using plotly to create an interactive bar plot of the year of release vs count of men as the older character

```{r}
plot_oldermen_byyear = ggplot(year_vs_oldermen, aes(x=release_year, y=n)) + 
  geom_bar(stat = "identity") + ggtitle("Count of older men characters by year")
plot_oldermen_byyear

#creating an interactive bar plot using ggplotly
interactive_plot = ggplotly(plot_oldermen_byyear)
interactive_plot
```

# Summary of findings part 2

According to the data from [Hollywood Age Gap](https://hollywoodagegap.com)

-   The movies included in the data have more men who are older characters when compared to women

-   By stratifying the movies by release year, it appears that the trend of men starring as the older character in the movie couple began to steadily increase in the early-to-mid 1990's and peaked in the year 2008, when 42 movies were released with an older male character in the movie couple

-   Since 2016, there appears to be a decrease in movies that have a man playing the older character in a couple

-   In 2022, there were just three movies released with a man playing he older character in the movie couple

# The Future 

Has Hollywood moved on from old man lovers? I wouldn't mind this. Either way, I am curious to see where Hollywood lovers will take us.
